---
title: Coryne BGC Alignments 
author: Rachel Myers
date: November 2022
output: 
    html_document:
        toc: true
        self.contained: true
---

# Coryne genomes analysis
# Preliminaries

## Credits
Author: Rachel Myers

Investigator: Dr. Matthew Kelly

Date: `r Sys.Date()`

## Recreating this report

1. Ensure that the R packages rmarkdown, knitr, and those listed below are installed

2. Ensure that the file Coryne_BGC_Alignments.rmd is in your working directory in R, change your R's working directory if needed 

3. Copy and paste the below code chunk into R, updating the path for the BGC fastq files

```{r recreate, echo = TRUE, eval = FALSE,  results = "asis"}

rm(list= ls(all= TRUE))
require("rmarkdown")
require("knitr")
setwd("Code/antismash")
fastadir <- "Data/BGC/MultipleAlignments/"
render("Coryne_BGC_Alignments_NAPAA.rmd", c("html_document"))

```

Random seed is sepcified for reproducibility of mulitple alignments and other random procesess


```{r loadpackages, echo = F, warnings = F, error = F, message = F}
require(muscle)
require(ggmsa)
require(msa)
require(seqinr)
require(ggtree)
require(ggdendro)
require(ggplot2)
set.seed(1234)
```

## Load protocore info 
```{r loadPCInfo, echo = FALSE, results = "asis"}
pcinfo <- read.csv("../../Data/BGC/MultipleAlignments/UniqueSequences.csv")

```

## Fastq file processing

* List all fasta files found in the fasta firectory noted above 

* For each file, generate a MUSCLE alignment 

```{r listfasta, echo = FALSE, results = "asis", message = FALSE, warning = FALSE}
# listfasta

fastafh = list.files(fastadir, pattern = "fasta", full.names = TRUE )
# reduce to the bacteriocin and RiPPS

fastafh = fastafh[grepl("NAPAA", fastafh)]


fastadata = lapply(fastafh, readAAStringSet )
names(fastadata) = gsub("respir_", "", gsub(".fasta", "", basename(fastafh)))

alnmts = lapply(fastadata, function(x) muscle(x, quiet = TRUE) )
names(alnmts) = gsub("respir_", "", gsub(".fasta", "", basename(fastafh)))

```

# NAPAA proto_gene sequence alignments

## NAPAA - All 

```{r poc_1a, echo = FALSE, results = "asis",fig.width = 8, fig.height =20,warning = FALSE, message = FALSE }

aln <- alnmts[["NAPAA"]]
aln1 <- msaConvert(aln, "seqinr::alignment")
writeXStringSet(as(aln, "AAStringSet"), file="aln.fasta")
```
```{r poc_2, echo = FALSE, results = "asis",fig.width = 14, fig.height =8 }

# visulaize the dendogram
cat("<br>Tree using hamming string distances based on the protein sequences<br>")
sdist <- stringDist(as(aln,"AAStringSet"), method="hamming")
clust <- hclust(sdist, method = "single")
dendr <- dendro_data(clust, type="rectangle") 
ggplot() + 
  geom_segment(data=segment(dendr), aes(x=x, y=y, xend=xend, yend=yend)) + 
  geom_text(data=dendr$labels, aes(x=x, y=y, label=label, hjust=-0.01), size=2.5) +
  coord_flip() + scale_y_reverse(expand=c(0.4, 0)) + 
  theme_dendro()
```

```{r poc_6, echo = FALSE, results = "asis", echo = FALSE, fig.width = 20, fig.height =15,warning = FALSE, message = FALSE}

cat("<br>Tree using Fitch similarity for amino acid sequences")
d <- dist.alignment(aln1, "similarity")
tr <- hclust(d)
#https://4va.github.io/biodatasci/r-ggtree.html
msaplot(p=ggtree(tr), fasta = "aln.fasta", offset =.3, width = 4) + geom_tiplab( hjust = -0.1)
```


## NAPAA- non Contig Edges 

```{r c2dsb_nce_1a, echo = FALSE, results = "asis",fig.width = 20, fig.height =6,warning = FALSE, message = FALSE }

aln <- alnmts[["NAPAAno_CEs"]]
aln1 <- msaConvert(aln, "seqinr::alignment")
writeXStringSet(as(aln, "AAStringSet"), file="aln.fasta")
```
```{r c2dsb_nce_2, echo = FALSE, results = "asis",fig.width =12 , fig.height =8 }

# visualize the dendogram
cat("<br>Tree using hamming string distances based on the protein sequences<br>")
sdist <- stringDist(as(aln,"AAStringSet"), method="hamming")
clust <- hclust(sdist, method = "single")
dendr <- dendro_data(clust, type="rectangle") 
ggplot() + 
  geom_segment(data=segment(dendr), aes(x=x, y=y, xend=xend, yend=yend)) + 
  geom_text(data=dendr$labels, aes(x=x, y=y, label=label, hjust=-0.01), size=2.5) +
  coord_flip() + scale_y_reverse(expand=c(0.4, 0)) + 
  theme_dendro()
```

```{r c2dsb_nce_6, echo = FALSE, results = "asis", echo = FALSE, fig.width = 20, fig.height =10,warning = FALSE, message = FALSE}

cat("<br>Tree using Fitch similarity for amino acid sequences")
d <- dist.alignment(aln1, "similarity")
tr <- hclust(d)
#https://4va.github.io/biodatasci/r-ggtree.html
msaplot(p=ggtree(tr), fasta = "aln.fasta", offset =.3, width = 4) + geom_tiplab( hjust = -0.1)
```

## Down-selection within clades that have high similarity 

Using the tree based on Fitch similarity for amino acid sequences

```{r ds_setup, echo = FALSE, results = "asis", fig.width = 8, fig.height = 10}
cutheight = .1


dendr <- dendro_data(tr, type="rectangle") 
ggplot() + 
  geom_segment(data=segment(dendr), aes(x=x, y=y, xend=xend, yend=yend)) + 
  geom_text(data=dendr$labels, aes(x=x, y=y, label=label, hjust=-0.01), size=2.5) +
  coord_flip() + scale_y_reverse(expand=c(0.4, 0)) + 
  theme_dendro() + geom_hline(yintercept = cutheight, color = "red")

cat("<br>Cut the tree at height =", cutheight, "<br>")

ct = cutree(tr, h = cutheight)
kable(table(ct))
```


## Alignment and tree using down-selected sequences
```{r ds_stage1, echo = FALSE, results = "asis", fig.width = 8, fig.height =6}
subsample <- data.frame(groupID = unique(ct),
                        nSequences = as.vector(table(ct)),
                        Sequences = sapply(unique(ct), function(x) paste( names(ct)[which(ct == x)], collapse = ";")),
                        SelectedID = sapply(unique(ct), function(x) sample( names(ct)[which(ct == x)], 1)),
                        stringsAsFactors = FALSE)
                        
fastatemp <-fastadata[["NAPAAno_CEs"]]

fastatemp <- fastatemp[which(names(fastatemp) %in% subsample$SelectedID), ]

temp_aln <- muscle(fastatemp , quiet = TRUE)
temp_aln1 <- msaConvert(temp_aln, "seqinr::alignment")
writeXStringSet(as(temp_aln, "AAStringSet"), file="aln.fasta")


cat("<br>Tree using Fitch similarity for amino acid sequences")
d <- dist.alignment(temp_aln1, "similarity")
tr <- hclust(d)
#https://4va.github.io/biodatasci/r-ggtree.html
msaplot(p=ggtree(tr), fasta = "aln.fasta", offset =1.3, width = 4) + geom_tiplab( hjust = -0.1)

NAPAA_downselection <- data.frame(ct)
NAPAA_downselection$Selected<- ifelse(row.names(NAPAA_downselection) %in% subsample$SelectedID, "Yes", "No")

NAPAA_downselection <- merge(NAPAA_downselection, pcinfo, by.x = "row.names", by.y = "SeqID", all.x = TRUE)

outfh = "../../Data/BGC/MultipleAlignments/NAPAA_list.csv"
write.csv(NAPAA_downselection, outfh, row.names = FALSE)

cat("<br>NAPAA down selection results saved to:", outfh)

```


# Session info
```{r sessionInfo}
sessionInfo()

```